---
title: "lauren_harris_snowshow_report"
author: "Lauren Harris"
date: "11/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(lubridate)
```

# Introduction

A 4 - 5 sentence "mini-introduction" to what this study is exploring, including basic background information about the subject, site and study (with citations as needed), and a brief overview of what is included in the exploratory report. You should probably do this last so you know what’s contained in the report

# Data and Analyses

A 3 - 4 sentence section in which you briefly describe the data, including citation & credit as needed, and provide an overview of the main analysis, statistical methods and tools used.

# Juvenile Hare Initial Investigation

## Annual Juvenile Hare Trap Counts
```{r}
# Read in data
hares <- read_csv(here("data", "bonanza_hares.csv"))

# reduce data set to juvenile hares only
juvenile_hares <- hares %>% 
  filter(age == "j") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date))

# plot the juvenile hares trapped per year
ggplot(data = juvenile_hares, aes(x = year)) +
  geom_bar() +
  labs(x = "Year", y = "Juvenile Hares Trapped", title = )
```


## Juvenile Hare Weights

## Juvenile Weight Comparison By Sex

## Relationship Between Juvenile Weight & Hind Foot Length

# Summary


# Citations