---
title: "lauren_harris_snowshoe_report"
author: "Lauren Harris"
date: "11/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(lubridate)
library(dplyr)
```

# Introduction

A 4 - 5 sentence "mini-introduction" to what this study is exploring, including basic background information about the subject, site and study (with citations as needed), and a brief overview of what is included in the exploratory report. You should probably do this last so you know what’s contained in the report

# Data and Analyses

A 3 - 4 sentence section in which you briefly describe the data, including citation & credit as needed, and provide an overview of the main analysis, statistical methods and tools used.

# Juvenile Hare Initial Investigation

## Annual Juvenile Hare Trap Counts
```{r}
# Read in data
hares <- read_csv(here("data", "bonanza_hares.csv"))

# reduce data set to juvenile hares only
juvenile_hares <- hares %>% 
  filter(age == "j") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date))

# plot the juvenile hares trapped per year
ggplot(data = juvenile_hares, aes(x = year)) +
  geom_bar() +
  labs(x = "Year", y = "Juvenile Hares Trapped") +
  theme_minimal()
```

**Figure 1.** Annual juvenile hare trap counts. Data comes from the Bonanza Creek Experimental Forest LTER site. Years range from `r min(juvenile_hares$year)` to `r max(juvenile_hares$year)`. 

```{r}
juvenile_hares_summary <- juvenile_hares %>% 
  group_by(year) %>% 
  summarize(count = n())

juvenile_hares_stats <- juvenile_hares_summary %>% 
  summarize(
    mean = mean(count), 
    median = median(count), 
    )
```

The minimum count of juvenile hares trapped for a given year was 0 during both 2002 and 2009, and the maximum was `r max(juvenile_hares_summary$count)` trapped which was during `r juvenile_hares_summary[juvenile_hares_summary$count == max(juvenile_hares_summary$count), ]$year`. The median number of juvenile hares trapped annually was `r juvenile_hares_stats$median` hares and the mean was `r juvenile_hares_stats$mean` hares. One notable trend in the data is that the number of juvenile hares trapped annually generally decreases per year. I would suggest that the amount of traps set be included in this dataset, then the number of juveniles trapped could be divided by the number of traps set. Then, the absolute counts would become success rates, which are more standardized, informative, and reflective of true hare populations. 

## Juvenile Hare Weights

## Juvenile Weight Comparison By Sex

## Relationship Between Juvenile Weight & Hind Foot Length

# Summary


# Citations
