---
title: "Snowshoe report"
author: "Lauren Harris"
date: "11/4/2021"
output: 
  html_document: 
    toc: yes
    number_sections: yes
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(lubridate)
library(dplyr)
library(kableExtra)
library(effsize)
```

# Introduction

A 4 - 5 sentence "mini-introduction" to what this study is exploring, including basic background information about the subject, site and study (with citations as needed), and a brief overview of what is included in the exploratory report. You should probably do this last so you know what’s contained in the report

# Data and Analyses

A 3 - 4 sentence section in which you briefly describe the data, including citation & credit as needed, and provide an overview of the main analysis, statistical methods and tools used.

# Juvenile Hare Initial Investigation

## Annual Juvenile Hare Trap Counts
```{r}
# read in data
hares <- read_csv(here("data", "bonanza_hares.csv"))

# reduce data set to juvenile hares only and translate dates into years
juvenile_hares <- hares %>% 
  filter(age == "j") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date))

ggplot(data = juvenile_hares, aes(x = year)) +
  geom_bar(fill = "darkslategray") +
  labs(x = "Year", y = "Juvenile Hares Trapped") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_x_continuous(breaks = seq(from = 1999, to = 2012, by = 1))
  
```

**Figure 1.** Annual juvenile hare trap counts. Annual counts of trapped juvenile hares from the Bonanza Creek Experimental Forest LTER site. (`r min(juvenile_hares$year)`-`r max(juvenile_hares$year)`). Data: Kielland et al. 2017. 

```{r}
# find counts of juvenile hares by year
juvenile_hares_summary <- juvenile_hares %>% 
  group_by(year) %>% 
  summarize(count = n())

# find mean and median counts of juvenile hares trapped 
juvenile_hares_stats <- juvenile_hares_summary %>% 
  summarize(
    mean = mean(count), 
    median = median(count))
```

### Summary

The minimum count of juvenile hares trapped for a given year was 0 (during 2002 and 2009), and the maximum was `r max(juvenile_hares_summary$count)` hares trapped, during `r juvenile_hares_summary[juvenile_hares_summary$count == max(juvenile_hares_summary$count), ]$year`. The median number of juvenile hares trapped annually was `r juvenile_hares_stats$median` hares and the mean was `r juvenile_hares_stats$mean` hares. One notable trend in the data is that the number of juvenile hares trapped annually generally decreased per year. Moving forward, I would suggest that the number of traps set be standardized across years, and recorded as part of the dataset. This would mean effort between years is the same, so the number of juvenile hares trapped is not impacted by sampling effort. Then, the data would more accurately reflect true hare populations. 

## Juvenile Hare Weights

```{r}
# group juvenile hares by sex and site, then get sample sizes and weights, then make sex and grid variables presentable
juv_hares_sex_site <- juvenile_hares %>% 
  group_by(sex, grid) %>% 
  mutate(sex = case_when(
    sex %in% "f" ~ "Female", 
    sex %in% "m" ~ "Male", 
    sex %in% NA ~ "Unknown")) %>% 
  mutate(grid = case_when(
    grid %in% "bonbs" ~ "Bonanza Spruce",
    grid %in% "bonmat" ~ "Bonanza Mature",
    grid %in% "bonrip" ~ "Bonanza Riparian"
  ))

ggplot(data = juv_hares_sex_site, aes(x = sex, y = weight, color = sex)) +
  geom_boxplot(outlier.color = "black") +
  scale_color_manual(values = c("lightpink2", "skyblue", "#D9B6F7")) +
  labs(y = "Body Weight (g)", x = "Sex") +
  facet_wrap(~grid) +
  geom_point(shape = 1, color = "darkgray") +
  theme(axis.text.x = element_text(angle = 90)) +
  theme(legend.position = "none") +
  stat_summary(fun = mean, geom ="point", shape = 5, size = 2, color = "black")

```

**Figure 2.** Distribution of body weights (g) for juvenile hares from Bonanza Creek Experimental Forest LTER site (`r min(juvenile_hares$year)`-`r max(juvenile_hares$year)`). Hares are grouped by sex and site. Individual data points are shown with hollow points and outliers are denoted by black points. The center lines in the boxes are the medians and the black hollow diamonds are means. The whiskers show minimums and maximums (not including outliers). Box endpoints are the 25th and 75th percentiles. Data: Kielland et al. 2017. 

**Table 1.** Summary statistics of weights of juvenile hares from Bonanza Creek Experimental Forest LTER site (`r min(juvenile_hares$year)`-`r max(juvenile_hares$year)`). Hares are grouped by sex and site. Statistics include mean (g), standard deviation (g), and sample size. No live hares of unknown sex were trapped at Bonanza Spruce so values there are NaN and NA. Data: Kielland et al. 2017.  

```{r}
# find mean, standard deviation and sample size for weights of juvenile hares
juv_hares_sex_site_stats <- juv_hares_sex_site %>% 
  group_by(sex, grid) %>% 
	summarise(mean = mean(weight, na.rm = TRUE),
	          median = median(weight, na.rm = TRUE),
	          sd = sd(weight, na.rm = TRUE),
	          n = n()) 

# make table of mean, median, standard deviation and sample size for juvenile hare weights by sex and site
juv_hares_sex_site_stats %>% 
  kable(col.names = c("Sex", "Site", "Mean (g)", "Median (g)", "Standard Deviation (g)", "Sample Size"), digits = 2) %>% 
  kable_styling(full_width = FALSE) %>% 
  column_spec(1:6, color = "black") %>% 
  row_spec(4:6, background = "lightgray")
```

### Summary

Mean and median body weights of juvenile hares are higher for males than females at all three sites. Juvenile hares with unknown sex only had recorded body weights at Bonanza Mature and Bonanza Riparian. For Bonanza Riparian, juvenile hares with unknown sex had much smaller mean and median weights compared to males and females at this site. Bonanza Spruce males and females had higher means and medians compared to all other juvenile hares (male, female, and unknown) at the other two sites. 

## Juvenile Weight Comparison By Sex

**Table 2.** Weights of juvenile hares from Bonanza Creek Experimental Forest LTER site (`r min(juvenile_hares$year)`-`r max(juvenile_hares$year)`) grouped by sex. Summary statistics include mean (g), standard deviation (g), and sample size. Data: Kielland et al. 2017. 

```{r}
# group juvenile hares by sex then find mean, standard deviaiton and sample size for weights
juv_sex_stats <- juvenile_hares %>% 
  group_by(sex) %>% 
	summarise(mean = mean(weight, na.rm = TRUE),
	          sd = sd(weight, na.rm = TRUE),
	          n = n()) %>% 
  filter(sex %in% c("m", "f"))

# make sex values presentable then make a table
juv_sex_stats %>% 
  mutate(sex = case_when(
    sex %in% "f" ~ "Female", 
    sex %in% "m" ~ "Male")) %>%
  kable(col.names = c("Sex", "Mean (g)", "Standard Deviation (g)", "Sample Size"), digits = 2) %>% 
  kable_styling(full_width = FALSE)

```

```{r}
# subset data for juvenile hares by sex
juv_hares_f <- juvenile_hares %>% 
  filter(sex == "f") %>% 
  select(weight)

juv_hares_m <- juvenile_hares %>% 
  filter(sex == "m") %>% 
  select(weight)

# find mean weights for males and females (for in-line code referencing)
f_juv_mean <- mean(juv_hares_f$weight, na.rm = TRUE)
m_juv_mean <- mean(juv_hares_m$weight, na.rm = TRUE)
```

```{r}
# create dataframe with males and females together
juv_hares_f_m <- juvenile_hares %>%
  filter(sex == "m" | sex == "f") %>% 
  select(sex, weight)

# exploratory tests before running t-test:

# histograms
ggplot(data = juv_hares_f_m, aes(x = weight))+
  geom_bar() +
  facet_wrap(~sex)

# qq plots
ggplot(data = juv_hares_f_m, aes(sample = weight))+
  geom_qq() +
  facet_wrap(~sex)

```

```{r}
# subset data with males and females but use pull() so I can run tests on it
juv_hares_f_pull <- juvenile_hares %>% 
  filter(sex == "f") %>% 
  pull(weight) %>% 
  na.omit()

juv_hares_m_pull <- juvenile_hares %>% 
  filter(sex == "m") %>% 
  pull(weight) %>% 
  na.omit()

# effect size with cohen's d
juv_hares_cohen <- cohen.d(juv_hares_m_pull, juv_hares_f_pull)

# t test, two samples but unpaired
juv_hares_ttest <- t.test(juv_hares_m_pull, juv_hares_f_pull)
```

### Summary

The mean weight of juvenile female hares is `r round(f_juv_mean, 2)` g, which is `r round((1-(f_juv_mean/m_juv_mean))*100, 2)`% smaller than the mean weight of juvenile male hares at `r round(m_juv_mean, 2)` g. The actual difference in mean weights is `r round(m_juv_mean - f_juv_mean, 2)` g. Based on a Cohen's d test, the effect of sex on body weight is not very large (d = `r round(juv_hares_cohen$estimate, 3)`) so the means are relatively close together. However, a t-test (Unpaired `r juv_hares_ttest$method`: p-value = `r round(juv_hares_ttest$p.value, 3)`, t(`r round(juv_hares_ttest$parameter, 2)`) = `r round(juv_hares_ttest$statistic, 2)`) shows that the effect is significant. This t-test purported a `r round((juv_hares_ttest$p.value)*100, 2)`% chance of finding means at least `r round(m_juv_mean - f_juv_mean, 2)` g different, if they are drawn from populations with the same mean. Diagnostic plots (histograms, qq plots) showed data and residuals were normally distributed. 

## Relationship Between Juvenile Weight & Hind Foot Length

```{r}
# data set with juvenile hares' weights and foot lengths 
juv_weight_foot <- juvenile_hares %>% 
  select(weight, hindft)

# plotting the relationship between juvenile weight & hind foot length
ggplot(data = juv_weight_foot, aes(x = hindft, y = weight)) +
  geom_point() +
  labs(x = "Hind Foot Length (mm)", y = "Body Weight (g)")
```

**Figure 3.** Relationship of hind foot length (mm) and body weight (g) of juvenile hares from Bonanza Creek Experimental Forest LTER site (`r min(juvenile_hares$year)`-`r max(juvenile_hares$year)`). Data: Kielland et al. 2017. 

```{r}
# trying out a linear model because boss said so
juv_hares_lm <- lm(weight ~ hindft, data = juv_weight_foot)
summary(juv_hares_lm)
```

# Summary


# Citations
